Open, privacy-preserving protocols for lawful surveillance

### 2016-08-03 paper/wpes16/chaining.tex
WPES submission edit pass done, down to 4 pages with good formatting

Without mechanisms to preserve privacy, a contact chaining search can collect a surprisingly large group of users' information. For example, if the average 
cell-phone user contacts 30 individuals within the period of the investigation,
a contact-chaining search out to distance three would capture 27,000 users on average - or many more if a heavy phone user is swept up by the search. In such a large group, the vast majority of contacts will not be collaborators of the the targeted, primary suspect in the investigation. These untargeted users may nevertheless face government scrutiny, intrusive investigation, or a risk that their sensitive communications histories may be leaked accidentally.

Despite this risk, 
we recognize the potential law-enforce\-ment value of information about the social contacts of targeted invidivuals. Therefore, we propose a \emph{lawful contact chaining protocol}. This protocol permits multiple government agencies working together to provide oversight and accountability, as advocated in \cite{sff-foci2014}. Our protocol focuses on the case in which the government seeks information from multiple telecommunications providers about the communication graph formed by phone calls and text messages. Using this protocol, the agencies can retrieve a set of encrypted records from multiple telecoms, {\it each of which holds only part of a larger communication graph}. This set of encrypted data 
contains the identities of users within a certain distance of a target, but the identities cannot be decrypted unless the agencies cooperate. Under the lawful processes we propose, this cooperation would take the form of intersecton with other sets of encrypted data. These sets can come from privacy-preserving contact chaining, from cell-tower dumps, or from other sources of information about suspects. Importantly, while any set may contain encrypted data about many untargeted users, very few users will appear in \emph{all} the sets, and those few users will be suitable targets for further lawful investigation.

The same principles of oversight and accountability provided by multiple government agencies can apply to contact chaining searches in other types of communication graphs, such as the social network graph of Twitter or Facebook. These cases do not require our protocol, however, since if one provider knows the entire communication graph, it can compute the output of the protocol without any interaction.

\subsection{Protocols For Privacy-Preserving Contact Chaining}

@@

We can assume that business phone numbers have many more contacts than personal phone numbers do. In most cases, knowing that two individuals have contacted the same business does not indicate that those individuals have a personal relationship. Therefore, the warrant also includes $d$, an upper bound on the degree of users that the agencies are willing to ``chain through.'' If a phone number has more than $d$ contacts, then the agencies do not follow paths to other users through that phone number in their search (but do include that number in the output). The agencies disregard $d$ for the initial target $x$, however, because they have already determined that contacts of $x$ are of potential interest, even if $x$ is a business.

@@

Every phone number in this set must be encrypted with each of the agencies' public ElGamal keys. The agencies should all have the same output.

The telecoms should not learn the agency's output. Instead, each telecom's output should contain only a list of which of the phone numbers it serves were sent to the government agencies. This allows the telecoms to play an additional accountability role.

With the encryptions of these phone numbers, the agencies can act as appropriate to further investigate them. In particular, the set of encrypted phone numbers can be intersected with, say, the encrypted numbers of people on a terrorist
watch list~\cite{sff-foci2014}.

@@

The agencies will represent their output as a list $\mathbf{C}$, containing agency ciphertexts. Each telecom $T$ will represent its output as a list $\mathbf{L}_T$, listing plaintext users served by that telecom whose information the agencies requested.  The protocol is as follows:

@@

\item Otherwise, $T(a)$ encrypts each neighbor $b$ of $a$ under the public key of $T(b)$, creating a telecom ciphertext for $b$.

@@

\item If $\de(a)>d$ and $j\neq k$ (i.e. $a\neq x$), the agencies discard all telecom ciphertexts received for $a$'s neighbors (i.e., agencies refuse to sign these ciphertexts in future steps of the protocol, and do not send them on to the telecoms).

@@

Note that the inner loop can be executed many times in parallel, up to the point of completely emptying $\mathbf{Q}$ at the beginning of the loop. Many messages to the same telecom can also be batched and sent together, thereby reducing the number of signing and verifying operations so that they depend only on $k$ and not on the size of the input or output.

Correctness and privacy of the basic protocol are argued 
in~\cite[Section 4.2]{sff-TR}.

\subsection{Performance of Privacy-Preserving Contact Chaining Protocol}

We implemented the basic protocol in Java and tested the implementation's 
running time, CPU time used, and data sent over the network. 
Our implementation uses the variant of our protocol in which the agencies 
completely exhaust the search queue $\mathbf{Q}$ each round, 
sending all queries at any given distance from $x$ to the 
telecoms at once in batches. This variant allows for greater parallelism. 
All of the telecoms receive their batch of queries at the same time and 
operate on those queries using eight parallel threads of computation.

@@

For our underlying communication graph, we used an a\-no\-ny\-mized data set of 1.6 million users from Pokec, a Slovakian social network~\cite{snapnets}. To replicate the multi-provider environment, we assigned each user to one of four telecom servers. The telecoms were each given a different number of the users, in proportion to the subscriber base of the largest four telecoms in the world~\cite{mobiforge}.

These results are important in evaluating how practical our lawful contact-chaining protocol would be it were put into practice by government agencies and telecoms. However, our data set is relatively small compared to the data\-ba\-ses held by real telecommunications companies, and each company handles that data using different technologies. The absolute running time and CPU usage of executing this protocol could vary from telecom to telecom. Therefore, we also produced a implementation of the contact-chaining protocol which omits all cryptographic operations. This version of the protocol does not preserve the privacy of users. By comparing the performance of our lawful contact-chaining protocol with the zero-cryptography contact-chaining protocol, however, we can get a sense of the ``cost of privacy and accountability.''

@@

%NOTE: Adding blank lines to this environment will interfere with the formatting.
\begin{figure*}[t]
\centering
\begin{subfigure}{0.66\columnwidth}
\centering
\begin{tikzpicture}[scale=0.66]
\begin{loglogaxis}[
	title={Protocol Runtime},
	title style={font=\Large,},
	xlabel={Ciphertexts in Result},
	ylabel={Running Time [min]},
	ylabel style={overlay},
	ticklabel style = {font=\large},
	xmin=10, xmax=1000000,
	ymin=0.01, ymax=100,
	%xtick={10, 100, 1000, 10000, 100000},
	%ytick={0.01, 0.1, 1, 10, 100, 1000},
	%log ticks with fixed point,
	legend pos=north west,
	ymajorgrids=true,
	xmajorgrids=true,
	grid style=dashed,
	legend style={cells={anchor=west}},
]
\addplot[
	color=blue,
	mark=o,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=RunningTime, col sep=comma] {chainingrsa.csv};	
\addplot[
	color=red,
	mark=square,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=RunningTime, col sep=comma] {chainingnocrypto.csv};
\legend{Lawful contact-chaining, Zero-crypto}
\end{loglogaxis}
\end{tikzpicture}
%\captionsetup{justification=centering}
%\caption{Protocol Runtime}
\label{fig:runtime}
\end{subfigure}\hfill
\begin{subfigure}{0.66\columnwidth}
\centering
\begin{tikzpicture}[scale=0.66]
\begin{loglogaxis}[
	title={All-Telecom CPU Time},
	title style={font=\Large,},
	xlabel={Ciphertexts in Result},
	ylabel={Total CPU Time [min]},
	ylabel style={overlay, anchor=north,},
	ticklabel style = {font=\large},
	xmin=10, xmax=1000000,
	ymin=0.1, ymax=1000,
	%xtick={10, 100, 1000, 10000, 100000},
	%ytick={0.01, 0.1, 1, 10, 100, 1000},
	%log ticks with fixed point,
	legend pos=north west,
	ymajorgrids=true,
	xmajorgrids=true,
	grid style=dashed,
	legend style={cells={anchor=west}},
]
\addplot[
	color=blue,
	mark=o,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=CPUTime, col sep=comma] {chainingrsa.csv};
\addplot[
	color=red,
	mark=square,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=CPUTime, col sep=comma] {chainingnocrypto.csv};
\legend{Lawful contact-chaining, Zero-crypto}
\end{loglogaxis}
\end{tikzpicture}
%\captionsetup{justification=centering}
%\caption{CPU Time}
\label{fig:cputime}
\end{subfigure}\hfill
\begin{subfigure}{0.66\columnwidth}
\centering
\begin{tikzpicture}[scale=0.66]
\begin{loglogaxis}[
	title={Data transferred},
	title style={font=\Large,},
	xlabel={Ciphertexts in Result},
	ylabel={Data transferred [MB]},
	ylabel style={overlay, anchor=north,},
	ticklabel style = {font=\large},
	xmin=10, xmax=1000000,
	ymin=0.1, ymax=10000,
	%xtick={10, 100, 1000, 10000, 100000},
	%ytick={0.01, 0.1, 1, 10, 100, 1000},
	%log ticks with fixed point,
	legend pos=north west,
	ymajorgrids=true,
	xmajorgrids=true,
	grid style=dashed,
	legend style={cells={anchor=west}},
]
\addplot[
	color=blue,
	mark=o,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=Bytes, col sep=comma] {chainingrsa.csv};	
\addplot[
	color=red,
	mark=square,
	only marks,
	mark size=1.5pt,
	]
	table [x=Ciphertexts, y=Bytes, col sep=comma] {chainingnocrypto.csv};
\legend{Lawful contact-chaining, Zero-crypto}
\end{loglogaxis}
\end{tikzpicture}
%\captionsetup{justification=centering}
%\caption{Network Data}
\label{fig:data}
\end{subfigure}
\captionsetup{justification=centering}
\caption{Performance of Lawful Contact-Chaining}
\label{fig:performance}
\end{figure*}

Our implementation of lawful contact-chaining performed well. Our experiments showed a linear relationship between the number of ciphertexts in the output and the running time, CPU time, and data usage of the protocol. We display graphs of our recorded data in Figure~\ref{fig:performance}. 
%Taking the average of all cases with $d>25$, the telecoms used 58.2 ms of CPU time per ciphertext. The agencies used, again in the average case, 2.0 ms of CPU time per ciphertext. Note that these times are the sums taken over all telecoms and all agencies respectively. Because the agencies have do very little cryptography in this protocol, we focus on the telecoms' CPU time in our evaluation.

We found that our protocol was able to process, in the average case, 197.4 ciphertexts per second. To return to our example from earlier of a network with an average of 30 contacts per user, a lawful contact-chaining search with $k=2$ would have 900 users in the output, and a search with $k=3$ would have 27,000 users in the output. To compare these times to some of our actual experiments, we found that a search that returned 937 ciphertexts took 6.86 seconds to run, and a search that returned 27,338 ciphertexts took 109.55 seconds to run. 
%To provide another point of comparison, recall that, in~\cite{sff-foci2014}, we studied the ``High Country Bandits'' case, in which the FBI performed an intersection of 150,000 phone number to help solve a series of bank robberies. In one of our experiments with lawful contact chaining, we find that a similarly sized data set of 149,535 ciphertexts took 625.08 seconds - 10.4 minutes - to compile with our protocol. Given the context of a criminal investigation, we feel these running times are quite reasonable.

@@

More detailed performance analysis is presented in~\cite[Section 4.3]{sff-TR}.

>>>>

Without mechanisms to preserve privacy, a contact-chaining search can collect a surprisingly large group of users' information. For example, if the average 
cell-phone user contacts 30 individuals within the period of the investigation,
a contact-chaining search out to distance three would capture 27,000 users on
average -- or many more if a heavy phone user is swept up by the search. In
such a large group, the vast majority of contacts will not be collaborators of
the targeted, primary suspect in the investigation. These untargeted users
may nevertheless face government scrutiny, intrusive investigation, or a risk
that their sensitive communications histories may be leaked accidentally.

Despite this risk, 
we recognize the potential law-enforce\-ment value of information about the
social contacts of targeted invidivuals. Therefore, we propose a \emph{lawful
contact-chaining protocol}. Such a protocol permits multiple government agencies
working together to provide oversight and accountability~\cite{sff-foci2014}.
Our protocol focuses on the case in which the government seeks information from
multiple telecommunications providers about the communication graph formed by
phone calls and text messages. Using this protocol, the agencies can retrieve a
set of encrypted records from multiple telecoms, {\it each of which holds only
part of a larger communication graph}. This set of encrypted data 
contains the identities of users within a certain distance of a target, but the
identities cannot be decrypted unless the agencies cooperate. Under the lawful
processes we propose, this cooperation would take the form of intersecton with
other sets of encrypted data. These sets can come from privacy-preserving
contact chaining, from cell-tower dumps, or from other sources of information
about suspects. While any set may contain encrypted data about
many untargeted users, few users will appear in \emph{all} the sets, and
those few will be suitable targets for further lawful investigation.

The same principles of oversight and accountability provided by multiple government agencies can apply to contact-chaining searches in other types of communication graphs, such as the social-network graph of Twitter or Facebook. These cases do not require our protocol, however: if one provider knows the entire communication graph, it can compute the output of the protocol without any interaction.

\subsection{Privacy-Preserving Contact Chaining}

@@

Business phone numbers often have many more contacts than personal
phone numbers do, and knowing that two individuals have contacted
the same business does not usually indicate
a relationship between those individuals.
Therefore, the warrant also includes $d$, an upper bound on the
degree of users that the agencies will ``chain through.'' If a phone
number has more than $d$ contacts, then the agencies do not follow paths to
other users through that phone number in their search (but do include that
number in the output). The agencies disregard $d$ for the initial target $x$,
however, because they have already determined that contacts of $x$ are of
potential interest, even if $x$ is a business.

@@

Every phone number in this set must be encrypted with each of the agencies' public ElGamal keys. The agencies should all have the same output.
The telecoms should not learn the agency's output. Instead, each telecom's output should contain only a list of which of the phone numbers it serves were sent to the government agencies. This allows the telecoms to play an additional accountability role.

The agencies can act as
appropriate to further investigate these encrypted phone numbers.
The set of encrypted
phone numbers can be intersected with, say, the encrypted numbers of people on
a terrorist
watch list~\cite{sff-foci2014}.

@@

The agencies represent their output as a list $\mathbf{C}$ of agency
ciphertexts. Each telecom $T$ represents its output as a list
$\mathbf{L}_T$ of plaintext users served by that telecom whose
information the agencies requested.  The protocol is as follows:

@@

\item Otherwise, $T(a)$ encrypts each neighbor $b$ of $a$ under $T(b)$'s public key, creating a telecom ciphertext for $b$.

@@

\item If $\de(a)>d$ and $j\neq k$ (i.e. $a\neq x$), the agencies discard all
telecom ciphertexts received for $a$'s neighbors (they refuse to sign
these ciphertexts in future protocol steps, and do not send them to
telecoms).

@@

The inner loop can be executed many times in parallel, up to the
point of completely emptying $\mathbf{Q}$ at the beginning of the loop. Many
messages to the same telecom can also be batched and sent together, thereby
reducing the number of signing and verifying operations so that they depend
only on $k$ and not on the size of the input or output.

Correctness and privacy of the basic protocol are argued 
in the longer technical report~\cite[Section 4.2]{sff-TR}.

\subsection{Contact-Chaining Protocol Performance}

\input{fig-chaining}

We implemented the basic protocol in Java and tested its
running time, CPU time, and network utilization.
Our implementation uses the variant of our protocol in which the agencies 
completely exhaust the search queue $\mathbf{Q}$ each round, 
sending all queries at any given distance from $x$ to the 
telecoms at once in batches, for greater parallelism. 
All telecoms receive their batch of queries at the same time and 
operate on those queries using eight parallel threads.

@@

For our contact graph, we used an anonymized data set of 1.6
million users from Pokec, a Slovakian social network~\cite{snapnets}. To
simulate multiple providers, we assigned each user to one of four
telecom servers.
Each telecom server was given a different number of the users,
in proportion to the subscriber bases
of the world's largest four telecoms~\cite{mobiforge}.

These results help in evaluating how practical our lawful
contact-chaining protocol would be in practice.
However, our data set is small compared to
the data\-ba\-ses held by real telecommunications companies, each company
handles that data using different technologies, and absolute costs
might vary.
Therefore, we also produced a implementation of the contact-chaining protocol
that omits all cryptographic operations, and does not
preserve the privacy of users. By comparing the performance of our lawful
contact-chaining protocol with the zero-cryptography protocol,
we can get a sense of the ``cost of privacy and accountability.''

@@

Our implementation of lawful contact-chaining performed well,
showing a linear relationship between the number of ciphertexts in the output
and the running time, CPU time, and data usage of the protocol. We display
graphs of our recorded data in Figure~\ref{fig:performance}. 
%Taking the average of all cases with $d>25$, the telecoms used 58.2 ms of CPU time per ciphertext. The agencies used, again in the average case, 2.0 ms of CPU time per ciphertext. Note that these times are the sums taken over all telecoms and all agencies respectively. Because the agencies have do very little cryptography in this protocol, we focus on the telecoms' CPU time in our evaluation.

We found that our protocol was able to process about 197
ciphertexts per second on average.
Returning to our example of a network with
an average of 30 contacts per user, a search with $k=2$
would have 900 users in the output, and a search with $k=3$ would have 27,000
users in the output. In our experiments,
we found that a search that returned 937 ciphertexts took 6.86 seconds to run,
and a search that returned 27,338 ciphertexts took 109.55 seconds to run.

%To provide another point of comparison, recall that, in~\cite{sff-foci2014}, we studied the ``High Country Bandits'' case, in which the FBI performed an intersection of 150,000 phone number to help solve a series of bank robberies. In one of our experiments with lawful contact chaining, we find that a similarly sized data set of 149,535 ciphertexts took 625.08 seconds - 10.4 minutes - to compile with our protocol. Given the context of a criminal investigation, we feel these running times are quite reasonable.

@@

More detailed analysis is presented in~\cite[Section 4.3]{sff-TR}.

### 2016-08-03 paper/wpes16/intro.tex
WPES submission edit pass done, down to 4 pages with good formatting

We now apply them to {\it contact chaining}.
The goal is to use the topology of a {\it communication 
graph} ({\it e.g.}, a phone-call graph, email graph, or social network) 
to identify associates (or ``contacts'') of 
lawfully targeted users~\cite{techdirt}. 
Government agencies can then investigate those associates to determine whether
they deserve further attention.
It is useful to consider both direct contacts, 
{\it i.e.}, users who are neighbors in the communication graph, and
extended contacts, {\it i.e.}, users who are at distance $k$ in the 
communication graph, for an appropriate constant $k$.
%In a phone-call graph, if Alice calls Bob, 
%and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and 
%Charlie), but Alice and Charlie are extended contacts (more precisely,
%contacts at distance 2).  
Without accountability and security mechanisms to
limit an investigation's scope, contact chaining in a mass-communication
network can sweep in a huge number of untargeted users.
In Section~\ref{sec-chaining}, we provide an accountable contact-chaining
protocol that bounds the scope of the search, uses encryption to protect 
untargeted users, and is computationally efficient (in that its time
complexity and communication complexity are both linear in the size of the 
output).

We posit that the
{\it Openness Principle} put forth in \cite{sff-foci2014} 
should govern all surveillance activity in a democracy.
\begin{compactenum}
\item[I]
Any surveillance or law-enforcement process
that obtains or uses private information
about untargeted users shall be
an open, public, unclassified process.
\item[II]
Any secret surveillance or law-enforcement processes shall use only:
\begin{compactenum}
\item[(a)] public information and
\item[(b)] private information about targeted users obtained under 
authorized warrants via open surveillance processes.
\end{compactenum}
\end{compactenum}
This principle can be viewed as a requirement that an open
``privacy firewall'' be placed between government agencies and citizens'
private information in a mass-communication network.  Processes that move
untargeted users' private information through the firewall must be open
processes.

>>>>

We now apply them to {\it contact chaining}.
The goal is to use the topology of a {\it communication 
graph} ({\it e.g.}, a phone-call graph, email graph, or social network) 
to identify associates (or ``contacts'') of 
lawfully targeted users~\cite{techdirt}. 
Agencies can then investigate those associates to determine whether
they deserve further attention.
It is useful to consider both direct contacts, 
{\it i.e.}, users who are neighbors in the communication graph, and
extended contacts, {\it i.e.}, users who are at distance $k$ in the 
communication graph, for an appropriate constant $k$.
%In a phone-call graph, if Alice calls Bob, 
%and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and 
%Charlie), but Alice and Charlie are extended contacts (more precisely,
%contacts at distance 2).  
Without accountability and security mechanisms to
limit an investigation's scope, contact chaining in a mass-communication
network can sweep in a huge number of untargeted users.
Section~\ref{sec-chaining} presents an accountable contact-chaining
protocol that bounds the scope of the search, uses encryption to protect 
untargeted users, and is computationally efficient,
with time and communication complexity linear in the size of the output.

We posit that the
{\it Openness Principle} put forth in \cite{sff-foci2014} 
should govern all surveillance activity in a democracy.
\begin{compactenum}
\item[I]
Any surveillance or law-enforcement process
that obtains or uses private information
about untargeted users shall be
an open, public, unclassified process.
\item[II]
Any secret surveillance or law-enforcement processes shall use only:
\begin{compactenum}
\item[(a)] public information and
\item[(b)] private information about targeted users obtained under 
authorized warrants via open processes.
\end{compactenum}
\end{compactenum}
This principle can be viewed as a requirement that an open
``privacy firewall'' be placed between government agencies and citizens'
private information in a mass-communication network.  Processes that move
untargeted users' private information through the firewall must be open
processes.
